---
title: "Object Recognition Dataset Analysis"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyr)
library(dplyr)
library(ggplot2)

library(splitstackshape)
set.seed(1)
```

``` {r, include=FALSE}
vocDataPath <- './analytics/voc_df.csv'
vgDataPath <- './analytics/vg_df.csv'
cocoDataPath <- './analytics/coco_df.csv'

vocData <- read.csv(vocDataPath)
vgData <- read.csv(vgDataPath)
cocoData <- read.csv(cocoDataPath)
```

```{r, include=FALSE}
vocFragment <- vocData %>%
  select(imageId = img, x, y, w, h, label = names ) %>%
  mutate(dataset = "VOC", imageId=as.character(imageId))

vgFragment <- vgData %>%
  select(imageId = img, x, y, w, h, label = names ) %>%
  mutate(dataset = "VG", imageId=as.character(imageId), x = x + w/2, y = y + h/2)

cocoFragment <- cocoData %>%
  select(imageId = image_id, x, y, w, h, label = name ) %>%
  mutate(dataset = "COCO", imageId=as.character(imageId), x = x + w/2, y = y + h/2)

globalData <- rbind(vocFragment, vgFragment, cocoFragment) %>%
  mutate(imageId = as.factor(imageId), dataset=as.factor(dataset))

```

# General statistics

```{r}
generalStats <- globalData %>%
  group_by(dataset) %>%
  summarize(images=n_distinct(imageId), objects=n(), labels=n_distinct(label), objPerImg=objects/images) %>%
  arrange(desc(images), objects)

knitr::kable(generalStats)
```

# Object Location

```{r}
globalData <- globalData %>%
  mutate(area = w * h)
```

```{r, fig.width=12, fig.height=5}
sampledData <- globalData %>%
  group_by(dataset) %>%
  sample_n(2000)

sampledData %>%
  select(dataset, x, y) %>%
  ggplot(aes(x = x, y = y, fill=dataset)) +
    facet_wrap(~ dataset) +
    geom_point(shape=21, alpha=0.2)
```

# Object Size

```{r, fig.width=12, fig.height=5}
sampledData %>%
  select(dataset, w, h, area) %>%
  ggplot(aes(x = w, y = h, fill=dataset)) +
    facet_wrap(~ dataset) +
    geom_point(shape=21, alpha=0.2)
```

# Number of objects per image

```{r, fig.width=12, fig.height=8}
globalData %>%
  group_by(dataset, imageId) %>%
  summarize(objects=n()) %>%
  ggplot(aes(x=dataset, y=objects, fill=dataset)) +
    geom_boxplot(alpha=0.7)
```

# Number of objects per image (descending) (Maybe better as a percentage of images...)

```{r, fig.width=12, fig.height=10}
objectsPerImage <- globalData %>%
  group_by(dataset, imageId) %>%
  summarize(objects=n()) %>%
  arrange(desc(objects)) %>%
  mutate(id=sequence(n())) %>%
  mutate(rank = rank(desc(objects))) %>%
  filter(id < 5000) 

objectsPerImage %>%
 ggplot(aes(x=id, y=objects, color=dataset)) +
   geom_smooth(stat='identity', alpha=0.7)
```

# Instances per label

```{r, fig.width=12, fig.height=8}
globalData %>%
  group_by(dataset, label) %>%
  summarize(objects=n()) %>%
  filter(label != "person") %>%
  ggplot(aes(x=dataset, y=objects, fill=dataset)) +
    geom_boxplot(alpha=0.7) +
    scale_y_log10()
```

# Number of instances per label (descending)

```{r, fig.width=12, fig.height=10}
instancesPerLabel <- globalData %>%
  group_by(dataset, label) %>%
  summarize(objects=n()) %>%
  arrange(desc(objects)) %>%
  mutate(id=sequence(n())) %>%
  mutate(rank = rank(desc(objects))) %>%
  filter(id < 100) 

instancesPerLabel %>%
 ggplot(aes(x=id, y=objects, color=dataset)) +
   geom_smooth(stat='identity', alpha=0.7) +
   scale_y_log10()
```

# Same as before as a percentage of the total images in the dataset (i.e.: normalized)


# Top 10 labels (VOC)

```{r}
top10 <- instancesPerLabel %>%
  filter(id < 10) %>%
  select(dataset, label, instances=objects) %>%
  arrange(dataset, desc(instances))
```

```{r}
knitr::kable(filter(top10, dataset=="VOC"))
```

# Top 10 labels (COCO)

```{r}
knitr::kable(filter(top10, dataset=="COCO"))
```

# Top 10 labels (VG)

```{r}
knitr::kable(filter(top10, dataset=="VG"))
```


# (Correlación entre objetos por imagen)


